---
title: "The Sleep-Activity Trade-Off (Pre-COVID Baseline)"
author: "by Lea N Harvin"
date: "2026-01-29"
output:
  github_document: default
  pdf_document: default
---

#### CSV resources: 
https://www.kaggle.com/datasets/arashnic/fitbit/data

https://raw.githubusercontent.com/GTPB/PSLS20/master/data/NHANES.csv

```{r}
#Install and load packages if needed
# install.packages('tidyverse')
library(dplyr)
library(tidyverse)
library(ggplot2)
```

#### Load CSV files
```{r}
# Create a dataframe named 'daily_activity'
daily_activity <- read.csv("raw_data/dailyActivity_merged.csv")

# Create dataframe for the sleep data. 
sleep_day <- read.csv("raw_data/sleepDay_merged.csv")

# Create dataframe for the weight log data. 
weight_log <- read.csv("raw_data/weightLogInfo_merged.csv")

# Create dataframe for NHANES Physical Activity data for comparison.
NHANES_activity2009_2011 <- read.csv("raw_data/NHANES.csv")
```


## Data Initial Exploration

**FitBit Daily Activity Data (Apr-May 2016)**
```{r}
# Identify all the columns in the daily_activity data.
colnames(daily_activity)
# Take a look at the daily_activity data.
head(daily_activity)
```

**FitBit Sleep Data**
```{r}
# Identify all the columns in the sleep_day data.
colnames(sleep_day)
# Take a look at the sleep_day data.
head(sleep_day)
```

**FitBit Weight Data**
```{r}
# Identify all the columns in the weight_log data.
colnames(weight_log)
# there is a BMI column in this dataframe you could compare to the NHANES data.
head(weight_log)

# Note that all 3 FitBit datasets have the 'Id' field - 
# this can be used to merge the datasets.
```

**NHANES Physical Activity Date (2009-2011)**
```{r}
# Identify all the columns in the NHANES_activity data.
colnames(NHANES_activity2009_2011)
head(NHANES_activity2009_2011)
```


#### Summary statistics 

How many unique participants are there in each dataframe? 

```{r}
# It looks like there may be more participants in the daily activity 
# dataset than the sleep dataset.
n_distinct(daily_activity$Id)             #  33
n_distinct(sleep_day$Id)                  #  24
n_distinct(weight_log$Id)                 #   8
n_distinct(NHANES_activity2009_2011$ID)   #6779
```
 How many observations are there in each dataframe?
 
```{r}
nrow(daily_activity)            #  940
nrow(sleep_day)                 #  413
nrow(weight_log)                #   67
nrow(NHANES_activity2009_2011)  #10000

# filter the original data frame to just the female participants
NHANES_females <- NHANES_activity2009_2011 %>% filter(Gender == "female")
nrow(NHANES_females)            #5020

```

##### For the daily activity dataframe:
```{r}
daily_activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes) %>%
  summary()

 #  TotalSteps    TotalDistance    SedentaryMinutes
 #  Min.   :    0   Min.   : 0.000   Min.   :   0.0  
 #  1st Qu.: 3790   1st Qu.: 2.620   1st Qu.: 729.8  
 #  Median : 7406   Median : 5.245   Median :1057.5  
 #  Mean   : 7638   Mean   : 5.490   Mean   : 991.2  
 #  3rd Qu.:10727   3rd Qu.: 7.713   3rd Qu.:1229.5  
 #  Max.   :36019   Max.   :28.030   Max.   :1440.0
```

##### For the sleep dataframe:
```{r}
# convert minutes of sleep to hours
sleep_day <- sleep_day %>%
  mutate(TotalHoursAsleep = TotalMinutesAsleep / 60)

sleep_day %>%  
  select(TotalSleepRecords,
         TotalHoursAsleep,
         TotalTimeInBed) %>%
  summary()

#TotalSleepRecords TotalHoursAsleep  TotalTimeInBed 
# Min.   :1.000     Min.   : 0.9667   Min.   : 61.0  
# 1st Qu.:1.000     1st Qu.: 6.0167   1st Qu.:403.0  
# Median :1.000     Median : 7.2167   Median :463.0  
# Mean   :1.119     Mean   : 6.9911   Mean   :458.6  
# 3rd Qu.:1.000     3rd Qu.: 8.1667   3rd Qu.:526.0  
# Max.   :3.000     Max.   :13.2667   Max.   :961.0
```

##### For the weight dataframe:
```{r}
weight_log %>%  
  select(WeightPounds,
         Fat,
         BMI) %>%
  summary()

#  WeightPounds        Fat             BMI       
# Min.   :116.0   Min.   :22.00   Min.   :21.45  
# 1st Qu.:135.4   1st Qu.:22.75   1st Qu.:23.96  
# Median :137.8   Median :23.50   Median :24.39  
# Mean   :158.8   Mean   :23.50   Mean   :25.19  
# 3rd Qu.:187.5   3rd Qu.:24.25   3rd Qu.:25.56  
# Max.   :294.3   Max.   :25.00   Max.   :47.54  
#                 NA's   :65 
```

##### For the NHANES dataframe (females only):
```{r}
NHANES_females %>%
  select(SleepHrsNight, 
         BMI) %>%
  summary()

#SleepHrsNight         BMI       
# Min.   : 2.000   Min.   :12.88  
# 1st Qu.: 6.000   1st Qu.:21.24  
# Median : 7.000   Median :25.56  
# Mean   : 6.991   Mean   :26.77  
# 3rd Qu.: 8.000   3rd Qu.:31.20  
# Max.   :12.000   Max.   :81.25  
# NA's   :1079     NA's   :179
```

What This Tells Us About the Sample's Activities:

```{r}
# Create a new column for sedentary hours --> this will be easier for audience to understand verses minutes.
daily_activity <- daily_activity %>%
  mutate(SedentaryHours = SedentaryMinutes / 60)

# Create the plot with hours instead of minutes
ggplot(data=daily_activity, aes(x=TotalSteps, y=SedentaryHours)) + 
  geom_point(alpha = 0.5) +  # Semi-transparent points
  geom_smooth(method = "lm", se = TRUE, color = "red") +  # Add trendline
  labs(title = "Relationship Between Daily Steps and Sedentary Time",
       x = "Total Steps",
       y = "Sedentary Hours",
       caption = "Data shows negative correlation: more steps = less sedentary time") +
  theme_minimal()

# Save plot
ggsave("Relationship_Between_Daily_Steps_and_Sedentary_Time.png", width = 8, height = 6)
```

What's the relationship between minutes asleep and time in bed?

```{r}
# ggplot(data=sleep_day, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + geom_point()

# convert minutes to hours
sleep_day <- sleep_day %>%
  mutate(TotalHoursAsleep = TotalMinutesAsleep / 60) %>%
  mutate(TotalHoursInBed = TotalTimeInBed / 60)

# add a trend line to the above; is it helpful to have or no?
ggplot(data=sleep_day, aes(x=TotalHoursAsleep, y=TotalHoursInBed)) + 
  geom_point() +
  geom_smooth(method = "lm", se = TRUE)+
  labs(title = "Sleep Efficiency: Hours Asleep vs. Hours in Bed",
       x = "Total Hours Asleep",
       y = "Total Hours in Bed")+
  theme_classic()

# Save plot
ggsave("Hours_Asleep_vs_Hours_in_Bed.png", width = 8, height = 6)
```

Calculate R² for Sleep Efficiency:

```{r}
# Calculate R² for Sleep Efficiency:
# Use a linear model (gives you more info)
sleep_model <- lm(TotalHoursInBed ~ TotalHoursAsleep, data = sleep_day)
summary(sleep_model)

# Extract R² from the model
model_r_squared <- summary(sleep_model)$r.squared
cat("\nR² from linear model:", round(model_r_squared, 3))
```

While there's a strong positive correlation between time in bed and actual 
sleep (R² = 0.866), the vertical distance between points and the trendline 
indicates time spent in bed but not sleeping.

This scatter plot below better reveals sleep efficiency among FitBit users. 

```{r}
ggplot(data=sleep_day, aes(x=TotalHoursAsleep, y=TotalHoursInBed)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +  # Perfect efficiency line
  labs(title = "Sleep Efficiency: Hours Asleep vs. Hours in Bed",
       x = "Total Hours Asleep",
       y = "Total Hours in Bed",
       caption = "Red dashed line = perfect sleep efficiency (time in bed = time asleep)\nBlue line = actual relationship showing sleep efficiency gap") +
  theme_minimal()

# Save plot
ggsave("Hours_Asleep_vs_Hours_in_Bed_with_efficiency.png", width = 8, height = 6)
```
What could these trends tell you about how to help market this product? Or areas where you might want to explore further?

- Develop/enhance **sleep efficiency metrics** that show users their actual sleep quality
- Create **sleep hygiene recommendations** to reduce time spent lying awake [I don't like the use of the word "hygiene"]
- Send **personalized insights**: 'You spent 45 minutes in bed but not asleep last night - try these tips to fall asleep faster'


Merging the two data sets together
```{r}
combined_data <- merge(sleep_day, daily_activity, by="Id") # yields 24 unique IDs

# change script above as outer-join
#combined_data <- merge(sleep_day, daily_activity, by="Id", all=TRUE)
# using the outer-join, yields 33 unique ids
```

Take a look at how many participants are in this data set.
```{r}
n_distinct(combined_data$Id)

# [1] 24
```


```{r}
# Extract just the date part (everything before the space) as a new column to compare on
sleep_day <- sleep_day %>%
  mutate(SleepActivityDate = substr(SleepDay, 1, nchar(SleepDay) - 12))  # Removes " 12:00:00 AM"

# Now merge on dates
combined_data_by_Id_and_Date <- merge(sleep_day, daily_activity, 
                       by.x=c("Id", "SleepActivityDate"), 
                       by.y=c("Id", "ActivityDate"))

```

Check data
```{r}
# check dimensions
nrow(sleep_day)                        # Original sleep records
nrow(daily_activity)                   # Original activity records  
nrow(combined_data_by_Id_and_Date)     # Merged records (should be less than or equal to smallest dataset)

# verify no duplicates were created
combined_data_by_Id_and_Date %>%
  group_by(Id, SleepActivityDate) %>%
  filter(n() > 1) %>%
  nrow()  # should return 0

# show sample
head(combined_data_by_Id_and_Date %>% 
  select(Id, SleepActivityDate, TotalHoursAsleep, TotalSteps, SedentaryMinutes))

#[1] 413
#[1] 940
#[1] 413
#[1] 6
```

Investigate the duplicates found.
```{r}
# view the actual duplicate records
duplicates <- combined_data_by_Id_and_Date %>%
  group_by(Id, SleepActivityDate) %>%
  filter(n() > 1) %>%
  arrange(Id, SleepActivityDate)

View(duplicates)
```

Was this duplicated before the merge?
```{r}
# check if sleep_day has duplicates BEFORE the merge
sleep_day %>%
  group_by(Id, SleepActivityDate) %>%
  filter(n() > 1) %>%
  nrow()

# check if daily_activity has duplicates
daily_activity %>%
  group_by(Id, ActivityDate) %>%
  filter(n() > 1) %>%
  nrow()

# [1] 6
# [1] 0
```

Duplicates identified in the original data. Now what?
```{r}
# remove exact duplicates
sleep_day_deduplicated <- sleep_day %>%
  distinct(Id, SleepActivityDate, .keep_all = TRUE)

# then merge
combined_data_by_Id_and_Date <- merge(sleep_day_deduplicated, daily_activity, 
                       by.x=c("Id", "SleepActivityDate"), 
                       by.y=c("Id", "ActivityDate"))
```

Check data again
```{r}
# check dimensions
nrow(sleep_day)                        # Original sleep records
nrow(daily_activity)                   # Original activity records  
nrow(combined_data_by_Id_and_Date)     # Merged records (should be less than or equal to smallest data set)

# verify no duplicates were created
combined_data_by_Id_and_Date %>%
  group_by(Id, SleepActivityDate) %>%
  filter(n() > 1) %>%
  nrow()  # should return 0

# show sample
head(combined_data_by_Id_and_Date %>% 
  select(Id, SleepActivityDate, TotalHoursAsleep, TotalSteps, SedentaryMinutes))

#[1] 413
#[1] 940
#[1] 410
#[1] 0
```

Let's explore this combined data.
```{r}
# Summary of combined data
summary(combined_data_by_Id_and_Date %>% 
  select(TotalHoursAsleep, TotalSteps, SedentaryMinutes, Calories))

# How many unique participants in the combined dataset?
length(unique(combined_data_by_Id_and_Date$Id))

# Average records per participant
410 / length(unique(combined_data_by_Id_and_Date$Id))


# TotalHoursAsleep    TotalSteps    SedentaryMinutes    Calories   
# Min.   : 0.9667   Min.   :   17   Min.   :   0.0   Min.   : 257  
# 1st Qu.: 6.0167   1st Qu.: 5189   1st Qu.: 631.2   1st Qu.:1841  
# Median : 7.2083   Median : 8913   Median : 717.0   Median :2207  
# Mean   : 6.9862   Mean   : 8515   Mean   : 712.1   Mean   :2389  
# 3rd Qu.: 8.1667   3rd Qu.:11370   3rd Qu.: 782.8   3rd Qu.:2920  
# Max.   :13.2667   Max.   :22770   Max.   :1265.0   Max.   :4900  
# [1] 24
# [1] 17.08333
```

What This Means for Your Bellabeat Case Study?

```{r}
# Scatter plot: Sleep vs Steps
ggplot(data=combined_data_by_Id_and_Date, aes(x=TotalHoursAsleep, y=TotalSteps)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(title = "Relationship Between Sleep Duration and Daily Steps",
       x = "Total Hours Asleep",
       y = "Total Steps",
       caption = "Data from 410 matched daily FitBit records") +
  theme_minimal()

# Save plot
ggsave("Sleep_Duration_and_Daily_Steps.png", width = 8, height = 6)
```
From the scatter plot:

  - Clear downward trend line
  - People sleeping 8+ hours average around 6,000-8,000 steps
  - People sleeping <6 hours average around 9,000-12,000 steps
  - Wide variation exists, but the pattern is consistent

Calculate the correlation
```{r}
# Calculate correlation coefficient
correlation <- cor(combined_data_by_Id_and_Date$TotalHoursAsleep, 
                   combined_data_by_Id_and_Date$TotalSteps, 
                   use = "complete.obs")

print(paste("Correlation between sleep and steps:", round(correlation, 3)))

# Statistical test
cor_test <- cor.test(combined_data_by_Id_and_Date$TotalHoursAsleep, 
                     combined_data_by_Id_and_Date$TotalSteps)
print(cor_test)


#[1] "Correlation between sleep and steps: -0.19"

#	Pearson's product-moment correlation

#data:  combined_data_by_Id_and_Date$TotalHoursAsleep and combined_data_by_Id_and_Date$TotalSteps
#t = -3.9164, df = 408, p-value = 0.0001054
#alternative hypothesis: true correlation is not equal to 0
#95 percent confidence interval:
# -0.28199288 -0.09525253
#sample estimates:
#       cor 
#-0.1903439
```

Statistical Results:

  - Correlation: -0.19 (weak to moderate negative correlation)
  - P-value: 0.0001054 (highly statistically significant - this is NOT random!)
  - Interpretation: People who sleep MORE tend to take FEWER steps, and vice versa

Take a look at different Sleep Groups

```{r}
# Create sleep categories
combined_data_by_Id_and_Date <- combined_data_by_Id_and_Date %>%
  mutate(SleepCategory = case_when(
    TotalHoursAsleep < 6 ~ "Poor Sleep (<6 hrs)",
    TotalHoursAsleep >= 6 & TotalHoursAsleep < 8 ~ "Moderate Sleep (6-8 hrs)",
    TotalHoursAsleep >= 8 ~ "Good Sleep (8+ hrs)"
  ))

# Compare average steps by sleep category
sleep_steps_summary <- combined_data_by_Id_and_Date %>%
  group_by(SleepCategory) %>%
  summarise(
    avg_steps = mean(TotalSteps),
    median_steps = median(TotalSteps),
    count = n()
  )

print(sleep_steps_summary)

# Visualize by category
ggplot(data=combined_data_by_Id_and_Date, aes(x=SleepCategory, y=TotalSteps, fill=SleepCategory)) +
  geom_boxplot() +
  labs(title = "Daily Steps by Sleep Duration Category",
       x = "Sleep Category",
       y = "Total Steps") +
  theme_minimal() +
  theme(legend.position = "none")

# Save plot
ggsave("Daily_Steps_by_Sleep_Duration_Category.png", width = 8, height = 6)


#SleepCategory             avg_steps   median_steps    count
#<chr>                     <dbl>       <dbl>           <int>
#Good Sleep (8+ hrs)	      7267.474	  6469.5	        116	
#Moderate Sleep (6-8 hrs)	8761.253	  9548.0	        194	
#Poor Sleep (<6 hrs)	      9484.030	  9814.5	        100	
```

From the boxplot:

  * Poor Sleep (<6 hrs): Median ~10,000 steps (HIGHEST activity!)
  * Moderate Sleep (6-8 hrs): Median ~9,500 steps
  * Good Sleep (8+ hrs): Median ~6,500 steps (LOWEST activity)


I'm not done yet. How does my combined data compare with the filtered NHANES 
data?

```{r}
# First, calculate summary stats for both datasets

# FitBit sleep summary (per person average)
fitbit_sleep_summary <- combined_data_by_Id_and_Date %>%
  group_by(Id) %>%
  summarise(avg_sleep = mean(TotalHoursAsleep, na.rm = TRUE)) %>%
  mutate(Source = "FitBit Sample (n=24)")

# NHANES sleep summary (females only, remove NAs)
nhanes_sleep_summary <- NHANES_females %>%
  filter(!is.na(SleepHrsNight)) %>%
  select(SleepHrsNight) %>%
  rename(avg_sleep = SleepHrsNight) %>%
  mutate(Source = "NHANES Female Population")

# Combine for comparison
sleep_comparison <- bind_rows(
  fitbit_sleep_summary %>% select(avg_sleep, Source),
  nhanes_sleep_summary
)

# Create comparison plot
ggplot(sleep_comparison, aes(x = Source, y = avg_sleep, fill = Source)) +
  geom_boxplot() +
  geom_hline(yintercept = 7, linetype = "dashed", color = "red") +
  geom_hline(yintercept = 8, linetype = "dashed", color = "blue") +
  labs(title = "Sleep Duration: FitBit Sample vs. NHANES Female Population",
       subtitle = "Red line = 7 hours (minimum recommended), Blue line = 8 hours",
       x = "",
       y = "Hours of Sleep per Night") +
  theme_minimal() +
  theme(legend.position = "none")

# Print summary statistics
cat("
FitBit Sleep Summary:
")
summary(fitbit_sleep_summary$avg_sleep)

cat("
NHANES Female Sleep Summary:
")
summary(nhanes_sleep_summary$avg_sleep)
```

```{r}
# I forgot to include BMI in the combined data set
# Create a date column in the same format as your other data
weight_log <- weight_log %>%
  mutate(WeightActivityDate = substr(Date, 1, nchar(Date) - 12))  # Remove time portion
```

Merge BMI into Combined Dataset
```{r}
# Merge weight/BMI data with your combined sleep/activity data
combined_data_with_BMI <- combined_data_by_Id_and_Date %>%
  left_join(weight_log %>% select(Id, WeightActivityDate, BMI), 
            by = c("Id" = "Id", "SleepActivityDate" = "WeightActivityDate"))

# Check how many records now have BMI
sum(!is.na(combined_data_with_BMI$BMI))

# View sample
head(combined_data_with_BMI %>% 
  select(Id, SleepActivityDate, TotalHoursAsleep, TotalSteps, BMI))
```

Handle low BMI Match Rate
```{r}
# Calculate average BMI per person
fitbit_person_summary <- combined_data_with_BMI %>%
  group_by(Id) %>%
  summarise(
    avg_sleep = mean(TotalHoursAsleep, na.rm = TRUE),
    avg_steps = mean(TotalSteps, na.rm = TRUE),
    avg_BMI = mean(BMI, na.rm = TRUE)
  ) %>%
  filter(!is.na(avg_BMI))  # Only keep people who have BMI data

# How many people have BMI data?
nrow(fitbit_person_summary)
```
Create the Comparisons
```{r}
# Use this plot in your analysis, it have better formatting.

# FitBit sleep (all people, not just those with BMI)
fitbit_sleep_all <- combined_data_by_Id_and_Date %>%
  group_by(Id) %>%
  summarise(avg_sleep = mean(TotalHoursAsleep, na.rm = TRUE)) %>%
  mutate(Source = "FitBit Sample")

# NHANES sleep
nhanes_sleep <- NHANES_females %>%
  filter(!is.na(SleepHrsNight)) %>%
  select(SleepHrsNight) %>%
  rename(avg_sleep = SleepHrsNight) %>%
  mutate(Source = "NHANES Females")

# Combine
sleep_comparison <- bind_rows(
  fitbit_sleep_all %>% select(avg_sleep, Source),
  nhanes_sleep
)

# Plot
ggplot(sleep_comparison, aes(x = Source, y = avg_sleep, fill = Source)) +
  geom_boxplot() +
  geom_hline(yintercept = 7, linetype = "dashed", color = "red", alpha = 0.7) +
  geom_hline(yintercept = 8, linetype = "dashed", color = "blue", alpha = 0.7) +
  labs(title = "Sleep Duration: FitBit Sample vs. NHANES Female Population",
       subtitle = "Red line = 7 hours (minimum), Blue line = 8 hours (optimal)",
       x = "",
       y = "Hours of Sleep per Night") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("FitBit Sample" = "#FF6B6B", "NHANES Females" = "#4ECDC4"))

# Summary statistics
cat("\n=== SLEEP COMPARISON ===\n")
cat("\nFitBit Sample (n =", nrow(fitbit_sleep_all), "people):\n")
summary(fitbit_sleep_all$avg_sleep)

cat("\nNHANES Females (n =", nrow(nhanes_sleep), "people):\n")
summary(nhanes_sleep$avg_sleep)

# Statistical test
t.test(fitbit_sleep_all$avg_sleep, nhanes_sleep$avg_sleep)

# Save plot
ggsave("sleepDuration_FitBitvsNHanes.png", width = 8, height = 6)
```

BMI Distribution
```{r}
#install.packages("viridis")
#library(viridis)
```

```{r}
# FitBit BMI (only those who logged weight)
fitbit_bmi <- fitbit_person_summary %>%
  filter(!is.na(avg_BMI)) %>%
  select(avg_BMI) %>%
  rename(BMI = avg_BMI) %>%
  mutate(Source = "FitBit Sample")

# NHANES BMI
nhanes_bmi <- NHANES_females %>%
  filter(!is.na(BMI)) %>%
  select(BMI) %>%
  mutate(Source = "NHANES Females")

# Combine
bmi_comparison <- bind_rows(fitbit_bmi, nhanes_bmi)

# Plot
ggplot(bmi_comparison, aes(x = Source, y = BMI, fill = Source)) +
  geom_boxplot() +
  geom_hline(yintercept = 25, linetype = "dashed", color = "black", alpha = 0.7) +
  geom_hline(yintercept = 18.5, linetype = "dashed", color = "gray40", alpha = 0.7) +
  labs(title = "BMI: FitBit Sample vs. NHANES Female Population",
       subtitle = "Gray line = 18.5 (underweight), Black line = 25 (overweight)",
       x = "",
       y = "Body Mass Index (BMI)") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("FitBit Sample" = "#9ecae1", "NHANES Females" = "#DE8F05"))

# Summary statistics
cat("\n=== BMI COMPARISON ===\n")
cat("\nFitBit Sample (n =", nrow(fitbit_bmi), "people):\n")
summary(fitbit_bmi$BMI)

cat("\nNHANES Females (n =", nrow(nhanes_bmi), "people):\n")
summary(nhanes_bmi$BMI)

# Statistical test
t.test(fitbit_bmi$BMI, nhanes_bmi$BMI)

# Save plot
ggsave("BMI_FitBitCombinedvsNHanes.png", width = 8, height = 6)
```
Combined Summary Table
```{r}
# Create comparison table
comparison_summary <- data.frame(
  Metric = c("Sleep Duration (hours)", "Sleep Duration (hours)", "Sleep Duration (hours)",
             "BMI", "BMI", "BMI",
             "Sample Size", "Sample Size"),
  Statistic = c("Median", "Mean", "Std Dev",
                "Median", "Mean", "Std Dev",
                "Participants", "Records"),
  FitBit = c(
    median(fitbit_sleep_all$avg_sleep),
    mean(fitbit_sleep_all$avg_sleep),
    sd(fitbit_sleep_all$avg_sleep),
    median(fitbit_bmi$BMI, na.rm = TRUE),
    mean(fitbit_bmi$BMI, na.rm = TRUE),
    sd(fitbit_bmi$BMI, na.rm = TRUE),
    nrow(fitbit_sleep_all),
    410
  ),
  NHANES = c(
    median(nhanes_sleep$avg_sleep),
    mean(nhanes_sleep$avg_sleep),
    sd(nhanes_sleep$avg_sleep),
    median(nhanes_bmi$BMI, na.rm = TRUE),
    mean(nhanes_bmi$BMI, na.rm = TRUE),
    sd(nhanes_bmi$BMI, na.rm = TRUE),
    nrow(NHANES_females),
    nrow(NHANES_females)
  )
)

# Round for readability
comparison_summary$FitBit <- round(comparison_summary$FitBit, 2)
comparison_summary$NHANES <- round(comparison_summary$NHANES, 2)

print(comparison_summary)

# Save as CSV for your report
write.csv(comparison_summary, "fitbit_nhanes_comparison.csv", row.names = FALSE)
```

